---
title: "Project"
author: "Lia Freed-Doerr"
date: "February 20, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(naniar)
```

```{r}
cdc <- readr::read_csv("cdc-113-ICD10.csv")
```

```{r}
clean_cdc <- cdc %>% 
  select(-c("State Code", "Year Code", "Ten-Year Age Groups", "Notes")) %>%
  na.omit() %>%
  na_if("Unreliable") %>%
  rename(crude_rate= "Crude Rate", 
         state= "State", 
         year= "Year",
         age_groups= "Ten-Year Age Groups Code", 
         cause_of_death= "ICD-10 113 Cause List",
         cause_of_death_code= "ICD-10 113 Cause List Code", 
         deaths= "Deaths", 
         population= "Population")
```

```{r}
clean_cdc %>% miss_var_summary
```

To do:
Get rid of hashtags, transform data to make code at end a separate column?
Frequency of deaths per each disease (compared to total cases disease, total deaths)
Explore: how many deaths reported per population?
Create a new crude rate column per 100,000
Look at different age groups, rate of deaths (Also would be a nice graph)
Across different states difference in deaths by certain diseases 
The rates of death by certain preventable illnesses--increasing or decreasing? (Graph)