---
title: "Project"
author: "Lia Freed-Doerr"
date: "February 20, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(naniar)
library(shiny)
library(shinydashboard)
```

```{r message=FALSE, warning=FALSE}
cdc <- readr::read_csv("cdc.csv")
cdc
```

```{r}
cdc %>% miss_var_summary
```

Get rid of hashtags

```{r}
new_cause_of_death <- str_remove(cdc$cause_of_death, "#")
cdc <- cdc %>%
  mutate(cause_of_death = new_cause_of_death)
cdc
```

```{r}
cdc$cause_of_death <- as.factor(cdc$cause_of_death)
cdc$cause_of_death_code <- as.factor(cdc$cause_of_death_code)
cdc$year <- as.factor(cdc$year)
cdc$race <- as.factor(cdc$race)
cdc$race_code <- as.factor(cdc$race_code)
cdc$gender <- as.factor(cdc$gender)
cdc$age_group <- as.factor(cdc$age_group)
cdc$deaths <- as.numeric(cdc$deaths)
cdc$population <- as.numeric(cdc$population)
cdc
```

Dawson: Death Frequency (as percent of population) for a Cause of Death by Year
```{r}
cdc %>% 
  filter(cause_of_death == "Falls (W00-W19)") %>% 

  group_by(year) %>% 
  summarise(
    population = sum(population, na.rm=T),
    #In the Data, N.S. age groups included in the population of all existing age groups. So it is OK to remove the NA values because it will product the same result.
    total_deaths = sum(deaths),
    percent= (total_deaths/population)*100
    ) %>% 
  ggplot(aes(x=year, y=percent, group=1))+geom_line()+labs(title="Falling Deaths in The United States", y="Percent of Population", x="Year")+
  theme(plot.title = element_text(size = 14, face = "bold", vjust = 3.5, hjust=.5, margin = margin(t=15, b=10)),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))
```



Expanding that into a shiny...

```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Number Deaths in CDC"),
  dashboardSidebar(),
  dashboardBody(
  fluidRow(
  box(title = "Plot Options", width = 3,
  selectInput("death", "Select Cause of Death", choices = c(levels(cdc$cause_of_death)),
              selected = "Falls (W00-W19)"),
  ), # close the first box
  box(title = "CDC Deaths", width = 7,
  plotOutput("plot", width = "800px", height = "500px")
  ) # close the second box
  ) # close the row
  ) # close the dashboard body
) # close the ui
server <- function(input, output, session) { 
  
  output$plot <- renderPlot({
    cdc %>% 
    filter(cause_of_death ==input$death) %>% 
    group_by(year) %>% 
  summarise(
    population = sum(population, na.rm=T),
    total_deaths = sum(deaths),
    percent= (total_deaths/population)*100
    ) %>% 
  ggplot(aes(x=year, y=percent, group=1))+geom_line()+labs(title="Deaths in The United States", y="Percent of Population", x="Year")+
      theme_light(base_size = 18) +
      theme(axis.text.x = element_text(angle = 0, hjust = 1))+
      labs(x = "Year", y = "Number")
  })
  
  # stop the app when we close it
  session$onSessionEnded(stopApp)
}  
shinyApp(ui, server)
```

To do:
Transform data to make code at end a separate column? What did I mean when I wrote this
Frequency of deaths per each disease (compared to total cases disease, total deaths)
Explore: how many deaths reported per population?
Create a new crude rate column per 100,000
Look at different age groups, rate of deaths (Also would be a nice graph)
Across different states difference in deaths by certain diseases 
The rates of death by certain preventable illnesses--increasing or decreasing? (Graph)

Make a shiny